---
title: "tmLQCD HMC profile"
subtitle: "`r args[2]`"
author: ""
date: "`r Sys.Date()`"
output:
  pdf_document: 
    keep_md: true 
    toc: true
  html_document: 
    toc: true
header-includes:
 - \usepackage{booktabs}
 - \usepackage{longtable}
 - \usepackage{array}
 - \usepackage{multirow}
 - \usepackage{wrapfig}
 - \usepackage{float}
 - \usepackage{colortbl}
 - \usepackage{pdflscape}
 - \usepackage{tabu}
 - \usepackage{threeparttable}
 - \usepackage{threeparttablex}
 - \usepackage[normalem]{ulem}
 - \usepackage{makecell}
 - \usepackage{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(knitr)
library(treemap)
library(hadron)
library(kableExtra)
library(dplyr)
```

```{r reading,echo=FALSE}
load("profile.RData")
```


\clearpage

# Total time per monomial 

```{r echo=FALSE, results="asis"}
plot_data <- rbind(total_per_mon,
                   data.frame(monomial = "other and unaccounted for",
                              time = total_time-sum(total_per_mon$time),
                              invocations = NA,
                              unit_time = NA,
                              prop = 100*(1 - sum(total_per_mon$time)/total_time)))
p <- ggplot2::ggplot(plot_data,
                     aes(x = 0, y = time, fill = reorder(monomial,prop))) +
     ggplot2::geom_bar(position = 'stack', stat = 'identity',
                       width = 1, colour = 'white') +
     ggrepel::geom_label_repel(aes(label = sprintf("%.1f%%",prop)),
                        position = position_stack(vjust = 0.5),
                        min.segment.length = unit(0, 'lines'),
                        colour = 'black', size = 4,
                        hjust = 0.5) +
     ggplot2::coord_polar('y', start = 0) +
     ggplot2::theme_bw() +
     ggplot2::labs(x = "", y = "") +
     ggplot2::theme(axis.text.y = element_blank(),
                    axis.ticks.y = element_blank())
plot(p)

cat(knitr::kable(plot_data,
                 format = 'latex',
                 booktabs = TRUE,
                 digits = 1,
                 caption = sprintf("Total time per monomial. Percentages are w.r.t. to the total time of %.1f seconds.",
                                   total_time)) %>%
    kableExtra::kable_styling(latex_options = "striped")
    )
```

\clearpage

# Hotspots by level

The total time spent in the benchmark was `r total_time` seconds. All percentages in the plot below are with respect to this total time.

```{r echo=FALSE, results="asis", warning = FALSE, message = FALSE}
plot_data <- dplyr::filter(data, level >= 1) %>%
             dplyr::group_by(name, level) %>%
             dplyr::summarise(time = sum(time)) %>%
             dplyr::ungroup() %>%
             dplyr::group_by(name) %>%
             dplyr::mutate(name = ifelse(time < 0.01*total_time, 'other', name)) %>%
             dplyr::ungroup() %>%
             dplyr::mutate(prop = 100 * time / total_time) %>%
             dplyr::mutate(label = sprintf("%s\n%.1f s (%.1f %%)", name, time, prop))

treemap(plot_data,
        index = c("level", "label"),
        vSize = "time",
        type = "index")
```

# Time by type and monomial

```{r echo=FALSE, results="asis"}
p <- ggplot2::ggplot(hb_acc_der_data,
                     aes(x = time, y = monomial, fill = type)) +
     ggplot2::geom_bar(position = 'stack', stat = 'identity',
                       width = 1.0, colour = 'white') +
     ggplot2::labs(fill = "", label = "", colour = "") +
     ggplot2::theme_bw()
plot(p)

cat("\\par\n")
cat(knitr::kable(dplyr::select(hb_acc_der_data, monomial, type, time, prop, invocations, unit_time),
                 format = 'latex',
                 booktabs = TRUE,
                 digits = 1,
                 linesep = "") %>%
    kableExtra::kable_styling(latex_options = "striped")
)
```

```{r echo=FALSE, results="asis"}
# Optional section if a QUDA profile was detected in the log file
if(!all(is.na(quda_data))){
  cat("\\clearpage")

  cat("\\section{overall QUDA profile}")

  cat(
    sprintf("The time spent in QUDA is %.1f seconds out of a total of %.1f seconds (%.1f %%).\\linebreak",
            sum(quda_data$time), total_time, 100*sum(quda_data$time)/total_time)
    )

  plot_data <- dplyr::mutate(quda_data, prop = 100*time/sum(time))
  p <- ggplot2::ggplot(plot_data,
                       aes(x = 0, y = time, fill = reorder(name,prop))) +
     ggplot2::geom_bar(position = 'stack', stat = 'identity',
                       width = 1, colour = 'white') +
     ggrepel::geom_label_repel(aes(label = sprintf("%.1f%%",prop)),
                               position = position_stack(vjust = 0.5),
                               min.segment.length = unit(0, 'lines'),
                               colour = 'black', size = 4,
                               hjust = 0.5,
                               max.overlaps = 20) +
     ggplot2::coord_polar('y', start = 0) +
     ggplot2::scale_fill_brewer(palette = "Set3") +
     ggplot2::theme_bw() +
     ggplot2::labs(x = "", y = "") +
     ggplot2::theme(axis.text.y = element_blank(),
                    axis.ticks.y = element_blank())
  plot(p)

  cat(knitr::kable(plot_data,
                   format = 'latex',
                   booktabs = TRUE,
                   digits = 1,
                   caption = sprintf("Time spent in different QUDA regions. Compare to total time of %.1f seconds.",
                                     total_time)) %>%
      kableExtra::kable_styling(latex_options = "striped")
     )
}
```

\clearpage

# Details by monomial and type

In the tables below, all percentages are given with respect to the given level. The monomials are ordered according to their contribution to the total time.

```{r echo=FALSE, results='asis'}
for( mon in total_per_mon$monomial ){
  cat(sprintf("\\subsection{%s}\n", hadron::escapeLatexSpecials(mon)))
  mon_data <- dplyr::filter(sum_data_per_mon, monomial == mon & level >= 2)
  for( tp in unique(mon_data$type) ){
    type_data <- dplyr::filter(mon_data, type == tp) %>%
                  dplyr::mutate(facet_label = sprintf("%s:%d", type, level))
    p <- ggplot2::ggplot(type_data, aes(y = time, x = "", fill = name)) +
         ggplot2::facet_wrap(vars(facet_label), ncol = 5, scales = "free") +
         ggplot2::geom_bar(position = 'stack', stat = 'identity',
                           width = 1.0, colour = 'white') +
         ggplot2::geom_label(aes(label = sprintf("%s %.1f %%", name, prop)),
                            position = position_stack(vjust = 0.5),
                            #min.segment.length = unit(0, 'lines'),
                            colour = 'white', size = 2,
                            hjust = 0.5) +
         ggplot2::ggtitle(sprintf("%s %s", mon, tp)) +
         ggplot2::theme_bw() +
         ggplot2::theme(legend.position = "none",
                        axis.text.x = element_blank(), 
                        axis.ticks.x = element_blank())
    plot(p)
    cat("\\par\n")
    cat(knitr::kable(dplyr::select(type_data, level, name, time, prop, invocations, unit_time),
                     format = 'latex',
                     booktabs = TRUE,
                     digits = 1,
                     linesep = "",
                     caption = sprintf("%s %s", mon, tp) ) %>%
          kable_styling(latex_options = "HOLD_position") %>%
        kableExtra::kable_styling(latex_options = c("striped"),
                                  stripe_index = which( (type_data$level %% 2) == 1 ))
    )
    cat("\\clearpage")
  }
}
```
